---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Rbduk

<!-- badges: start -->
<!-- badges: end -->

The goal of Rbduk is to provide BDUK analysts with a centralised set of repeatable functions for BDUK-specific R work.

## Installation

You can install the released version of Rbduk from [sam.atkin@dcms.gov.uk](sam.atkin@dcms.gov.uk) with:

``` r
devtools::install.github("NAME/Rbduk")
```

## is_integer64

Within GCP, large numbers (eg UPRNs) are stored are 64-bit integers. R, on the whole, does not like 64-bit integer, and whilst it can cope, it's easier to read in these columns are numeric, or convert them to numeric within R. To see if an object is currently a 64-bit integer, this function can be used.

```{r is_integer64}
library(Rbduk)

is_integer64(100)

is_integer64("x")

```

## pretty_postcode

This function allows us to enter a postcode in any format and any case, and convert it into pretty format XX(X)(X)(Y)XXX, where X is the postcode, and Y is the specified seperator, which is a space by default. The case can also be specified, and is upper case by default.

```{r pretty_postcode}
library(Rbduk)

pretty_postcode("SW1a2nP")

pretty_postcode("SW1a2nP", sep="")

pretty_postcode("SW1a2nP", sep=".")

pretty_postcode("SW1a2nP", sep=".", uppercase=FALSE)

```

